(working title) Local Physiological Adaptation in \textit{Mimulus cardinalis}

### 2015-04-29 ms/ms.tex
Worked on Discussion.

@@

As with previous studies in \textit{Mimulus} and many other species [CITES], latitude \textit{per se} predicted trait variation more than many obvious environmental variables (e.g. temperature or precipitation) that covary with latitude. We used Random Forest regression \citep{Liaw_Wiener_2002} to identify putative climatic factors underlying trait-latitude associations in \textit{M. cardinalis}. We did this by looking for overlap between climatic variables that best predict latitude of \textit{M. cadinalis} populations and climatic variables that best predict trait variation across our 16 focal populations. We used a representative set of 178 recent (since 2000) known \textit{M. cardinalis} occurences. These occurences were thinned by 50\% to correct for uneven sampling from a comprehensive set of herbarium records and an exhaustive field survey in 2010-11 \citep{Angert_ENM}. For each occurence, we used a 90m digital elevation model from HydroSHEDS \citep{Lehner_etal_2006} to extract elevation. Monthly interpolated climate layers were calculated using ClimateWNA \citep{Wang_etal_2012}, which accurately downscales climate data specifically for the rugged topography of western North America. For each occurence, we calculated bioclimatic variables using the biovars function in the R package \pkg{dismo} \citep{Hijmans_etal_2014}. In total, we included X environmental variables from ClimateWNA and 17? bioclimatic variables (Table~\ref{table:TableS_}). For each variable, we calculated both a 30-year normal by averging annual values between 1981 and 2010 and 30-year coefficient of variation, a standardized metric of interannual climatic variation). Temperatures were converted to Kelvin to be on an interval scale appropriate for calculating the coefficient of variation.

Selection at given site might not reflect simply the local environment, but rather the average environment in a surrounding region, since alleles will experience selection across populations linked by migration. We therefore calculated spatially averaged \cdm{maybe spatially-integrated is better term?} climate means and coefficients of variation. We sampled climate at 1000 random points (at 90-m resolition) within a 10 km buffer around focal populations. Since \textit{M. cardinalis} is found exclusively in riparian areas, we only selected points along streams using the National Hydrogeoraphy Dataset \citep{NHD}. Climatic means and CVs were weighted by the predicted suitability of each point, using ENM models from \cite{Angert_ENM}.

To identify climatic variables that were closely associated with latitude in \textit{M. cardinalis}, we used a "Variable Selection Using Random Forest" (VSURF) algorithm in the R \pkg{VSURF} version 0.8.2 \citep{Genuer_etal_2014}. In this algorithm, climatic variables are regressed against latitude, and the most important variables were selected after a three-stage selection process that progressively excludes unimportant variables.

Predictions: looked for climatic variables (means or CVs) that were strong predictors of both population latitudes (from all card pops) and trait variation in 16 focal popultions. If spatial average environment is important, then expect greater predictive power and closer association with latitude-predictors than point averages.

Evaluated fit on test set (correlation between predicted versus actual latitude) and on pseudoabsence set.
, Despite this common, and long-standing pattern, only certain traits In particular, adaptation to the local abiotic environment is generally thought to be the major cause of natural selection. Physiology, the study of organismal function, connects fitness to the abiotic environment, yet we actually know little about the physiological basis of adaptation to different environments within a species. To address this question, we looked at physiological variation across a broad latitudual gradient within \textit{Mimulus cardinalis}...

% Methods for spatial-average selection
For 16 focal populations, I took the average climate within a 10 km buffer, clipped to streams based on the National Hydrological Database \citep{NHD} weighted by climatic suitability determined [multimodel average/ensemble?] from ecological niche modeling \citep{Angert_ENM}.

@@

{\ttfamily\noindent\color{warningcolor}{\#\# Warning in VSURF.parallel.formula(linmod.PC1, data = focClim[, c("{}PC1"{}, : VSURF with a formula-type call outputs selected variables\\\#\# which are indices of the input matrix based on the formula:\\\#\# you may reorder these to get indices of the original data}}\begin{alltt}
  \hlcom{# Plot most important climatic predictor versus PC1}
  \hlkwd{pdf}\hlstd{(}\hlkwd{paste}\hlstd{(pathFig,} \hlstr{"/Figure_bio16cva_v_Lat.pdf"}\hlstd{,} \hlkwc{sep} \hlstd{=} \hlstr{""}\hlstd{),} \hlnum{5}\hlstd{,} \hlnum{5}\hlstd{)}
  \hlkwd{par}\hlstd{(}\hlkwc{mar} \hlstd{=} \hlkwd{c}\hlstd{(}\hlnum{5}\hlstd{,} \hlnum{5}\hlstd{,} \hlnum{1}\hlstd{,} \hlnum{1}\hlstd{))}
  \hlkwd{plot}\hlstd{(focClim}\hlopt{$}\hlstd{bio16_cva, focClim}\hlopt{$}\hlstd{PC1,} \hlkwc{xlim} \hlstd{=} \hlkwd{c}\hlstd{(}\hlnum{0.25}\hlstd{,} \hlnum{0.55}\hlstd{),} \hlkwc{ylim} \hlstd{=} \hlkwd{c}\hlstd{(}\hlopt{-}\hlnum{4}\hlstd{,} \hlnum{4}\hlstd{),} \hlkwc{axes} \hlstd{= F,}
    \hlkwc{frame.plot} \hlstd{= T,} \hlkwc{type} \hlstd{=} \hlstr{"n"}\hlstd{,} \hlkwc{xlab} \hlstd{=} \hlstr{""}\hlstd{,} \hlkwc{ylab} \hlstd{=} \hlstr{""}\hlstd{)}
  \hlkwd{title}\hlstd{(}\hlkwc{xlab} \hlstd{=} \hlstr{"Precipitation of Wettest Quarter"}\hlstd{,} \hlkwc{cex.lab} \hlstd{=} \hlnum{1.5}\hlstd{)}
  \hlkwd{title}\hlstd{(}\hlkwc{xlab} \hlstd{=} \hlstr{"(Coefficient of variation)"}\hlstd{,} \hlkwc{cex.lab} \hlstd{=} \hlnum{1}\hlstd{,} \hlkwc{line} \hlstd{=} \hlnum{4}\hlstd{)}
  \hlkwd{title}\hlstd{(}\hlkwc{ylab} \hlstd{=} \hlkwd{substitute}\hlstd{(}\hlkwd{plain}\hlstd{(Principal}\hlopt{~}\hlstd{Component}\hlopt{~}\hlnum{1}\hlstd{)}\hlopt{~}\hlkwd{group}\hlstd{(}\hlstr{"("}\hlstd{, x,} \hlstr{")"}\hlstd{),}
    \hlkwd{list}\hlstd{(}\hlkwc{x} \hlstd{= pve)))}
  \hlkwd{axis}\hlstd{(}\hlnum{1}\hlstd{,} \hlkwc{lwd} \hlstd{=} \hlnum{0}\hlstd{,} \hlkwc{lwd.ticks} \hlstd{=} \hlnum{1}\hlstd{)}
  \hlkwd{axis}\hlstd{(}\hlnum{2}\hlstd{,} \hlkwc{at} \hlstd{=} \hlkwd{seq}\hlstd{(}\hlopt{-}\hlnum{4}\hlstd{,} \hlnum{4}\hlstd{,} \hlnum{2}\hlstd{),} \hlkwc{las} \hlstd{=} \hlnum{1}\hlstd{,} \hlkwc{lwd} \hlstd{=} \hlnum{0}\hlstd{,} \hlkwc{lwd.ticks} \hlstd{=} \hlnum{1}\hlstd{)}
  \hlstd{fitBio16cva} \hlkwb{<-} \hlkwd{lm}\hlstd{(PC1} \hlopt{~} \hlstd{bio16_cva,} \hlkwc{data} \hlstd{= focClim)}

  \hlcom{# polygon of confidence intervals}
  \hlstd{x} \hlkwb{<-} \hlkwd{seq}\hlstd{(}\hlkwd{min}\hlstd{(focClim}\hlopt{$}\hlstd{bio16_cva),} \hlkwd{max}\hlstd{(focClim}\hlopt{$}\hlstd{bio16_cva),} \hlkwc{l} \hlstd{=} \hlnum{1e3}\hlstd{)}
  \hlstd{u} \hlkwb{<-} \hlkwd{predict}\hlstd{(fitBio16cva,} \hlkwc{new} \hlstd{=} \hlkwd{data.frame}\hlstd{(}\hlkwc{bio16_cva} \hlstd{= x))}
  \hlstd{s} \hlkwb{<-} \hlkwd{sqrt}\hlstd{(}\hlkwd{sapply}\hlstd{(x,} \hlkwa{function}\hlstd{(}\hlkwc{X}\hlstd{)} \hlkwd{t}\hlstd{(}\hlkwd{c}\hlstd{(}\hlnum{1}\hlstd{, X))} \hlopt{%*%} \hlkwd{vcov}\hlstd{(fitBio16cva)} \hlopt{%*%} \hlkwd{c}\hlstd{(}\hlnum{1}\hlstd{, X)))}
  \hlstd{confInt} \hlkwb{<-} \hlkwd{data.frame}\hlstd{(}\hlkwc{lowerCI} \hlstd{=} \hlkwd{qnorm}\hlstd{(}\hlnum{0.025}\hlstd{, u, s),} \hlkwc{upperCI} \hlstd{=} \hlkwd{qnorm}\hlstd{(}\hlnum{0.975}\hlstd{, u, s))}
  \hlkwd{polygon}\hlstd{(}\hlkwd{c}\hlstd{(x,} \hlkwd{rev}\hlstd{(x)),} \hlkwd{c}\hlstd{(confInt}\hlopt{$}\hlstd{lowerCI,} \hlkwd{rev}\hlstd{(confInt}\hlopt{$}\hlstd{upperCI)),}
    \hlkwc{col} \hlstd{=} \hlstr{"grey80"}\hlstd{,} \hlkwc{border} \hlstd{=} \hlnum{NA}\hlstd{)}
  \hlkwd{lines}\hlstd{(x, confInt}\hlopt{$}\hlstd{lowerCI,} \hlkwc{lwd} \hlstd{=} \hlnum{2}\hlstd{,} \hlkwc{lty} \hlstd{=} \hlnum{2}\hlstd{)}
  \hlkwd{lines}\hlstd{(x, confInt}\hlopt{$}\hlstd{upperCI,} \hlkwc{lwd} \hlstd{=} \hlnum{2}\hlstd{,} \hlkwc{lty} \hlstd{=} \hlnum{2}\hlstd{)}
  \hlkwd{lines}\hlstd{(}\hlkwd{range}\hlstd{(focClim}\hlopt{$}\hlstd{bio16_cva),} \hlkwd{range}\hlstd{(}\hlkwd{predict}\hlstd{(fitBio16cva)),} \hlkwc{lwd} \hlstd{=} \hlnum{2}\hlstd{)}
  \hlkwd{points}\hlstd{(focClim}\hlopt{$}\hlstd{bio16_cva, focClim}\hlopt{$}\hlstd{PC1,} \hlkwc{pch} \hlstd{=} \hlnum{21}\hlstd{,} \hlkwc{col} \hlstd{=} \hlstr{"black"}\hlstd{,} \hlkwc{bg} \hlstd{= primShade0,}
    \hlkwc{cex} \hlstd{=} \hlnum{1.5}\hlstd{,} \hlkwc{lwd} \hlstd{=} \hlnum{2}\hlstd{)}
  \hlkwd{dev.off}\hlstd{()}
\end{alltt}
\begin{verbatim}
## pdf 
##   2
\end{verbatim}
\begin{alltt}
  \hlcom{# Overlap between latitude predictors and phenotype predictors}
  \hlstd{imp2} \hlkwb{<-} \hlstd{climVars[vsurfPC1}\hlopt{$}\hlstd{varselect.pred]}
  \hlstd{nImpPC1} \hlkwb{<-} \hlkwd{length}\hlstd{(imp2)}

@@

\begin{table}[ht]
   \centering
   \topcaption{Climatic variables}
   \begin{tabular}{@{} ll @{}}
      \toprule
  Climate variable & Abbreviation \\
      \midrule
  % climateWNA VARIABLES:
	DD\_0	& degree-days below 0\celsius (chilling degree-days)  \\
	DD5	  & degree-days above 5\celsius (growing degree-days)   \\
	DD\_18	& degree-days below 18\celsius (heating degree-days)  \\
	DD18	& degree-days above 18\celsius (cooling degree-days)  \\
	NFFD	& number of frost-free days                           \\
	PAS	  & precipitation as snow (mm) between August in previous year and July in current  \\
	Eref	& Hargreaves reference evaporation (mm) \\
	CMD	  & Hargreaves climativ moisture deficit (mm) \\
	MAR   &	mean annual solar radiation (NOTE: removing because too many missing values) \\
	RH    &	mean annual relative humidity \\
  % BIOCLIM VARIABLES:
  bio1	&	Annual Mean Temperature	\\
	bio2	&	Mean Diurnal Range (Mean of monthly (max temp - min temp))	\\
	bio3	&	Isothermality (bio2/bio7) (* 100)	\\
	bio4	&	Temperature Seasonality (standard deviation *100)	\\
	bio5	&	Max Temperature of Warmest Month	\\
	bio6	&	Min Temperature of Coldest Month	\\
	bio7	&	Temperature Annual Range (bio5-bio6)	\\
	bio8	&	Mean Temperature of Wettest Quarter	\\
	bio9	&	Mean Temperature of Driest Quarter	\\
	bio10	&	Mean Temperature of Warmest Quarter	\\
	bio11	&	Mean Temperature of Coldest Quarter	\\
	bio12	&	Annual Precipitation	\\
	bio15	&	Precipitation Seasonality (Coefficient of Variation)	\\
	bio16	&	Precipitation of Wettest Quarter	\\
	bio17	&	Precipitation of Driest Quarter	\\
	bio18	&	Precipitation of Warmest Quarter	\\
	bio19	&	Precipitation of Coldest Quarter	\\
	    \bottomrule
   \end{tabular}
\end{table}

%--------------------------------------------------
% Results
%--------------------------------------------------

@@

\begin{table}[ht]
   \centering
   \topcaption{Population, treatment effects}
   \begin{tabular}{@{} llllllll @{}}
      \toprule
                                   & Germination & Leaf expansion & Stem elongation & Photosynthesis & Intrin. Photo & Mortality \\
                                   & survreg     & lmer           & lmer            & lmer           & lmer          & glm \\
      \midrule
  Population                       &
    ***  &
    *** &
    *** &
    *** & 
    *** & 
    *** & \\
  Temperature                      &
    NA                                                          &
    ***    &
    *** &
    *** & 
     &
    *** & \\
  Water                            &
    NA                                                          &
    ***   &
    ***   &
    *** & 
     &     
    *** & \\
  Pop $\times$ Temp                &
    NA                                                          &
     &
     &
    * & 
     &     
    * & \\
  Pop $\times$ Water               &
     NA                                                          &
    * &
     &
     & 
     &     
     & \\
  Temp $\times$ Water              &
    NA                                                          &
     &
    *** &
     & 
     &     
    *** & \\
  Pop $\times$ Temp $\times$ Water &
    NA                                                          &
     &
     &
     & 
     &     
     & \\
      \bottomrule
   \end{tabular}
   \label{table:Table_AnovaSummary}
\end{table}

@@

\subsection*{Latitude and phenotypic clines are associated with more variable precipitation}

Latitude explains variation
Presumably selective agent is not latitude, but something related to it
Exploratory approach to identify putative selective agent to test in future studies
We used a machine learning approach...identified interannual variation in precipitation (Fig~\ref{fig:Fig_ClimVarImp})
Obviously, cannot identify alternative variables that were not included in analysis.

@@

1. How does local adaptation occur? It's complex (gene flow, shape of tradeoffs, demography) and although there is abundant evidence for local adaptaiton, we do not a general sense for when local adaptation succeeds or fails. For this, we need a system in which there are axes of variation and constraint, traits with known physiological function?, and selective agents

@@

@@

\begin{figure}
	\centerline{\includegraphics[width=0.5\textwidth]{Figures/Figure_Height_Lat.pdf}}
	\usefont{T1}{phv}{m}{n}
	\fontsize{10}{12}
	\selectfont
	\caption[Southern populations grow faster (stem elongation rate).]{CAPTION}
	\label{fig:Fig_height}
\end{figure}

>>>>

% Toggle these commands to get version for doing word count
% \linenumbers
% \pagenumbering{gobble}
% \raggedright

@@

As with previous studies in \textit{Mimulus} and many other species [CITES, \citep{Huey_etal}, mosquitoes, what else?], latitude \textit{per se} predicted trait variation more than many obvious environmental variables (e.g. temperature or precipitation) that covary with latitude. We used Random Forest regression \citep{Liaw_Wiener_2002} to identify putative climatic factors underlying trait-latitude associations in \textit{M. cardinalis}. We did this by looking for overlap between climatic variables that best predict latitude of \textit{M. cadinalis} occurence records and climatic variables that best predict trait variation across our 16 focal populations. For brevity, we refer to these as Climate-Latitude and Climate-Trait variables. We selected Climate-Latitude and Climate-Trait variables independently using Random Forest (VSURF algorithm in the R \pkg{VSURF} version 0.8.2 \citep{Genuer_etal_2014}). From VSURF models, we kept only variables selected for prediction, the most stringent criteria.

@@

{\ttfamily\noindent\color{warningcolor}{\#\# Warning in VSURF.parallel.formula(linmod.PC1, data = focClim[, c("{}PC1"{}, : VSURF with a formula-type call outputs selected variables\\\#\# which are indices of the input matrix based on the formula:\\\#\# you may reorder these to get indices of the original data}}\begin{alltt}
  \hlstd{vsurfPC1_SA} \hlkwb{<-} \hlkwd{VSURF.parallel}\hlstd{(linmod.PC1,} \hlkwc{data} \hlstd{= focClimSA[,} \hlkwd{c}\hlstd{(}\hlstr{"PC1"}\hlstd{, climVars)],}
    \hlkwc{ncores} \hlstd{=} \hlnum{3}\hlstd{)}
\end{alltt}

{\ttfamily\noindent\color{warningcolor}{\#\# Warning in VSURF.parallel.formula(linmod.PC1, data = focClimSA[, c("{}PC1"{}, : VSURF with a formula-type call outputs selected variables\\\#\# which are indices of the input matrix based on the formula:\\\#\# you may reorder these to get indices of the original data}}\begin{alltt}
  \hlcom{# Plot most important climatic predictor versus PC1}
  \hlkwd{pdf}\hlstd{(}\hlkwd{paste}\hlstd{(pathFig,} \hlstr{"/Figure_bio16cva_v_Lat.pdf"}\hlstd{,} \hlkwc{sep} \hlstd{=} \hlstr{""}\hlstd{),} \hlnum{5}\hlstd{,} \hlnum{5}\hlstd{)}
  \hlkwd{par}\hlstd{(}\hlkwc{mar} \hlstd{=} \hlkwd{c}\hlstd{(}\hlnum{5}\hlstd{,} \hlnum{5}\hlstd{,} \hlnum{1}\hlstd{,} \hlnum{1}\hlstd{))}
  \hlkwd{plot}\hlstd{(focClim}\hlopt{$}\hlstd{bio16_cva, focClim}\hlopt{$}\hlstd{PC1,} \hlkwc{xlim} \hlstd{=} \hlkwd{c}\hlstd{(}\hlnum{0.25}\hlstd{,} \hlnum{0.55}\hlstd{),} \hlkwc{ylim} \hlstd{=} \hlkwd{c}\hlstd{(}\hlopt{-}\hlnum{4}\hlstd{,} \hlnum{4}\hlstd{),} \hlkwc{axes} \hlstd{= F,}
    \hlkwc{frame.plot} \hlstd{= T,} \hlkwc{type} \hlstd{=} \hlstr{"n"}\hlstd{,} \hlkwc{xlab} \hlstd{=} \hlstr{""}\hlstd{,} \hlkwc{ylab} \hlstd{=} \hlstr{""}\hlstd{)}
  \hlkwd{title}\hlstd{(}\hlkwc{xlab} \hlstd{=} \hlstr{"Precipitation of Wettest Quarter"}\hlstd{,} \hlkwc{cex.lab} \hlstd{=} \hlnum{1.5}\hlstd{)}
  \hlkwd{title}\hlstd{(}\hlkwc{xlab} \hlstd{=} \hlstr{"(Coefficient of variation)"}\hlstd{,} \hlkwc{cex.lab} \hlstd{=} \hlnum{1}\hlstd{,} \hlkwc{line} \hlstd{=} \hlnum{4}\hlstd{)}
  \hlkwd{title}\hlstd{(}\hlkwc{ylab} \hlstd{=} \hlkwd{substitute}\hlstd{(}\hlkwd{plain}\hlstd{(Principal}\hlopt{~}\hlstd{Component}\hlopt{~}\hlnum{1}\hlstd{)}\hlopt{~}\hlkwd{group}\hlstd{(}\hlstr{"("}\hlstd{, x,} \hlstr{")"}\hlstd{),}
    \hlkwd{list}\hlstd{(}\hlkwc{x} \hlstd{= pve)))}
  \hlkwd{axis}\hlstd{(}\hlnum{1}\hlstd{,} \hlkwc{lwd} \hlstd{=} \hlnum{0}\hlstd{,} \hlkwc{lwd.ticks} \hlstd{=} \hlnum{1}\hlstd{)}
  \hlkwd{axis}\hlstd{(}\hlnum{2}\hlstd{,} \hlkwc{at} \hlstd{=} \hlkwd{seq}\hlstd{(}\hlopt{-}\hlnum{4}\hlstd{,} \hlnum{4}\hlstd{,} \hlnum{2}\hlstd{),} \hlkwc{las} \hlstd{=} \hlnum{1}\hlstd{,} \hlkwc{lwd} \hlstd{=} \hlnum{0}\hlstd{,} \hlkwc{lwd.ticks} \hlstd{=} \hlnum{1}\hlstd{)}
  \hlstd{fitBio16cva} \hlkwb{<-} \hlkwd{lm}\hlstd{(PC1} \hlopt{~} \hlstd{bio16_cva,} \hlkwc{data} \hlstd{= focClim)}

  \hlcom{# polygon of confidence intervals}
  \hlstd{x} \hlkwb{<-} \hlkwd{seq}\hlstd{(}\hlkwd{min}\hlstd{(focClim}\hlopt{$}\hlstd{bio16_cva),} \hlkwd{max}\hlstd{(focClim}\hlopt{$}\hlstd{bio16_cva),} \hlkwc{l} \hlstd{=} \hlnum{1e3}\hlstd{)}
  \hlstd{u} \hlkwb{<-} \hlkwd{predict}\hlstd{(fitBio16cva,} \hlkwc{new} \hlstd{=} \hlkwd{data.frame}\hlstd{(}\hlkwc{bio16_cva} \hlstd{= x))}
  \hlstd{s} \hlkwb{<-} \hlkwd{sqrt}\hlstd{(}\hlkwd{sapply}\hlstd{(x,} \hlkwa{function}\hlstd{(}\hlkwc{X}\hlstd{)} \hlkwd{t}\hlstd{(}\hlkwd{c}\hlstd{(}\hlnum{1}\hlstd{, X))} \hlopt{%*%} \hlkwd{vcov}\hlstd{(fitBio16cva)} \hlopt{%*%} \hlkwd{c}\hlstd{(}\hlnum{1}\hlstd{, X)))}
  \hlstd{confInt} \hlkwb{<-} \hlkwd{data.frame}\hlstd{(}\hlkwc{lowerCI} \hlstd{=} \hlkwd{qnorm}\hlstd{(}\hlnum{0.025}\hlstd{, u, s),} \hlkwc{upperCI} \hlstd{=} \hlkwd{qnorm}\hlstd{(}\hlnum{0.975}\hlstd{, u, s))}
  \hlkwd{polygon}\hlstd{(}\hlkwd{c}\hlstd{(x,} \hlkwd{rev}\hlstd{(x)),} \hlkwd{c}\hlstd{(confInt}\hlopt{$}\hlstd{lowerCI,} \hlkwd{rev}\hlstd{(confInt}\hlopt{$}\hlstd{upperCI)),}
    \hlkwc{col} \hlstd{=} \hlstr{"grey80"}\hlstd{,} \hlkwc{border} \hlstd{=} \hlnum{NA}\hlstd{)}
  \hlkwd{lines}\hlstd{(x, confInt}\hlopt{$}\hlstd{lowerCI,} \hlkwc{lwd} \hlstd{=} \hlnum{2}\hlstd{,} \hlkwc{lty} \hlstd{=} \hlnum{2}\hlstd{)}
  \hlkwd{lines}\hlstd{(x, confInt}\hlopt{$}\hlstd{upperCI,} \hlkwc{lwd} \hlstd{=} \hlnum{2}\hlstd{,} \hlkwc{lty} \hlstd{=} \hlnum{2}\hlstd{)}
  \hlkwd{lines}\hlstd{(}\hlkwd{range}\hlstd{(focClim}\hlopt{$}\hlstd{bio16_cva),} \hlkwd{range}\hlstd{(}\hlkwd{predict}\hlstd{(fitBio16cva)),} \hlkwc{lwd} \hlstd{=} \hlnum{2}\hlstd{)}
  \hlkwd{points}\hlstd{(focClim}\hlopt{$}\hlstd{bio16_cva, focClim}\hlopt{$}\hlstd{PC1,} \hlkwc{pch} \hlstd{=} \hlnum{21}\hlstd{,} \hlkwc{col} \hlstd{=} \hlstr{"black"}\hlstd{,} \hlkwc{bg} \hlstd{= primShade0,}
    \hlkwc{cex} \hlstd{=} \hlnum{1.5}\hlstd{,} \hlkwc{lwd} \hlstd{=} \hlnum{2}\hlstd{)}
  \hlkwd{dev.off}\hlstd{()}
\end{alltt}
\begin{verbatim}
## pdf 
##   2
\end{verbatim}
\begin{alltt}
  \hlcom{# Overlap between latitude predictors and phenotype predictors}
\hlcom{#### NEED TO COMPARE RESULTS FROM POINT DATA AND SPATIALLY AVERAGED CLIMATE!}
  \hlstd{imp2} \hlkwb{<-} \hlstd{climVars[vsurfPC1}\hlopt{$}\hlstd{varselect.pred]}
  \hlcom{# imp2 <- climVars[vsurfPC1_SA$varselect.pred]}
  \hlstd{nImpPC1} \hlkwb{<-} \hlkwd{length}\hlstd{(imp2)}

@@

For these analyses, we compiled a representative set of 178 recent (since 2000) known \textit{M. cardinalis} occurences. These occurences were thinned by 50\% to correct for uneven sampling from a comprehensive set of herbarium records and an exhaustive field survey in 2010-11 \citep{Angert_ENM}. For each occurence, we used a 90m digital elevation model from HydroSHEDS \citep{Lehner_etal_2006} to extract elevation. Monthly interpolated climate layers were calculated using ClimateWNA \citep{Wang_etal_2012}, which accurately downscales climate data specifically for the rugged topography of western North America. For each occurence, we calculated bioclimatic variables using the biovars function in the R package \pkg{dismo} \citep{Hijmans_etal_2014}. In total, we included 24 cliamte variables, 9 from ClimateWNA and 15 bioclimatic variables (Table~\ref{table:TableS_ClimVars}). For each variable, we calculated both a 30-year normal by averging annual values between 1981 and 2010 and 30-year coefficient of variation, a standardized metric of interannual climatic variation. Temperatures were converted to Kelvin to be on a ratio scale appropriate for calculating the coefficient of variation. 
%--------------------------------------------------
% Results
%--------------------------------------------------

@@

\begin{table}[ht]
   \centering
   \topcaption{Population, treatment effects}
   \begin{tabular}{@{} llllllll @{}}
      \toprule
                                   & Germination & Leaf expansion & Shoot elongation & Photosynthesis & Intrin. Photo & Mortality \\
                                   & survreg     & lmer           & lmer            & lmer           & lmer          & glm \\
      \midrule
  Population                       &
    ***  &
    *** &
    *** &
    *** & 
    *** & 
    *** & \\
  Temperature                      &
    NA                                                          &
    ***    &
    *** &
    *** & 
     &
    *** & \\
  Water                            &
    NA                                                          &
    ***   &
    ***   &
    *** & 
     &     
    *** & \\
  Pop $\times$ Temp                &
    NA                                                          &
     &
     &
    * & 
     &     
    * & \\
  Pop $\times$ Water               &
     NA                                                          &
    * &
     &
     & 
     &     
     & \\
  Temp $\times$ Water              &
    NA                                                          &
     &
    *** &
     & 
     &     
    *** & \\
  Pop $\times$ Temp $\times$ Water &
    NA                                                          &
     &
     &
     & 
     &     
     & \\
      \bottomrule
   \end{tabular}
   \label{table:Table_AnovaSummary}
\end{table}

@@

\subsection*{Climatic variability best explains phenotypic divergence}

Latitudinal clines are common, but it is often difficult to ascribe this variation to a particular selective agent. For \textit{M. cardinalis}, interannual variation in precipitation over the past 30 years is very closely related to the latitude of recently-recorded occurences of this species (Fig.~\ref{fig:Fig_ClimVarImp}A). The two most important Climate-Latitude variables were also strong predictors of position along the fast-slow growth axis (Fig.~\ref{fig:Fig_ClimVarImp}B). This overlap between Climate-Latitude and Climate-Trait variables suggests that interannual variation in precipitation is an important selective agent in \textit{M. cardinalis}. Specifically, we hypothesize that more frequent droughts (greater precipitation cv) in Southern populations selects for an `annual-ized' life history, as we detail in the Discussion. We must qualify these results because our analysis obviously cannot rule out that alternative variables not included in the analysis may be more important.

@@

The observation that populations are locally adapted to their environment is old and well-documented in many species. New frontiers in local adaptation research seek a general explanation for why local adaptation often suceeds but sometimes fails, why certain traits respond to selection whereas others are constrained, and what environmental drivers are most important [CITES such as Kawecki]. Understanding why and how populations have adapted to different environments in the past is especially significant now as a scientific basis for predicting how species' ranges will shift under climate change [CITES, \citep{Catullo_etal_2015}]. In this study, we found evidence for one of two common signature of local adaptation, pointing to both phenotypic lability and constraint. In particular, our data suggest that photosynthesis and intrinsic growth rate vary among populations, whereas the fundamental abiotic niche is more conserved, and that climatic variation may be a more important selective agent than the average. Next, we tie these results into the broader threads of evolutionary theory in order to figure out what factors best explain our data.

As long as environments vary and there exists genetic variation in fitness, why shouldn't local adapted genotypes evolve? Evolutionary theory indicates that the shape of fitness tradeoffs, demography, and gene flow can constrain adaptation \citep{Levins_1968, Ronce_Kirkpatrick_2001}. In \textit{M. cardinalis} we found substantial genetic variation among populations along a phenotypic axis from fast to slow growth that varied over a large spatial scale (Fig.~\ref{fig:Fig_PC1vLat}). If this variation is adaptive, it could be maintained as long as fitness tradeoffs are not too Th... fitness tradeoffs are not too strong, demographic symmetry, and attenuation of gene flow.

@@

	\bottomrule
	\end{tabular}
	\end{center}
	\label{tables:TableS_InitialSize}
\end{table}

\begin{table}[ht]
   \centering
   \topcaption{Climatic variables}
   \begin{tabular}{@{} ll @{}}
      \toprule
  Climate variable & Abbreviation \\
      \midrule
  % climateWNA VARIABLES:
	DD\_0	& degree-days below 0\celsius (chilling degree-days)  \\
	DD5	  & degree-days above 5\celsius (growing degree-days)   \\
	DD\_18	& degree-days below 18\celsius (heating degree-days)  \\
	DD18	& degree-days above 18\celsius (cooling degree-days)  \\
	NFFD	& number of frost-free days                           \\
	PAS	  & precipitation as snow (mm) between August in previous year and July in current  \\
	Eref	& Hargreaves reference evaporation (mm) \\
	CMD	  & Hargreaves climativ moisture deficit (mm) \\
	MAR   &	mean annual solar radiation (NOTE: removing because too many missing values) \\
	RH    &	mean annual relative humidity \\
  % BIOCLIM VARIABLES:
  bio1	&	Annual Mean Temperature	\\
	bio2	&	Mean Diurnal Range (Mean of monthly (max temp - min temp))	\\
	bio3	&	Isothermality (bio2/bio7) (* 100)	\\
	bio4	&	Temperature Seasonality (standard deviation *100)	\\
	bio5	&	Max Temperature of Warmest Month	\\
	bio6	&	Min Temperature of Coldest Month	\\
	bio7	&	Temperature Annual Range (bio5-bio6)	\\
	bio8	&	Mean Temperature of Wettest Quarter	\\
	bio9	&	Mean Temperature of Driest Quarter	\\
	bio10	&	Mean Temperature of Warmest Quarter	\\
	bio11	&	Mean Temperature of Coldest Quarter	\\
	bio12	&	Annual Precipitation	\\
	bio15	&	Precipitation Seasonality (Coefficient of Variation)	\\
	bio16	&	Precipitation of Wettest Quarter	\\
	bio17	&	Precipitation of Driest Quarter	\\
	bio18	&	Precipitation of Warmest Quarter	\\
	bio19	&	Precipitation of Coldest Quarter	\\
	    \bottomrule
   \end{tabular}
   \label{table:TableS_ClimVars}
\end{table}

@@

\begin{figure}
	\centerline{\includegraphics[width=0.5\textwidth]{Figures/Figure_Height_Lat.pdf}}
	\usefont{T1}{phv}{m}{n}
	\fontsize{10}{12}
	\selectfont
	\caption[Southern populations grow faster (shoot elongation rate).]{CAPTION}
	\label{fig:Fig_height}
\end{figure}

\subsection*{Spatially-averaged climate}

% Methods for spatial-average selection
Selection at given site might not reflect simply the local environment, but rather the average environment in a surrounding region, since alleles will experience selection across populations linked by migration. We therefore calculated spatially averaged climate means and coefficients of variation. We sampled climate at 1000 random points (at 90-m resolition) within a 10 km buffer around focal populations. Since \textit{M. cardinalis} is found exclusively in riparian areas, we only selected points along streams using the National Hydrogeoraphy Dataset \citep{NHD}. Climatic means and CVs were weighted by their climatic suitability as determined using a multimodel ensemble average of ecological niche models \citep{Angert_ENM}.
